---
title: "Introduction to ausflora"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ausflora}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When working with biodiversity data, it is important to verify  taxonomic names with an authoritative list and correct any out-of-date names. The 'ausflora' package simplifies this process by:

-   Accessing up-to-date taxonomic information from the [Australian Plant Census](https://biodiversity.org.au/nsl/services/search/taxonomy) and the [Australia Plant Name Index](https://biodiversity.org.au/nsl/services/search/names).
-   Aligning authoritative names to your taxonomic names using our [algorithm]()
-   Updating your taxonomic names in a transparent manner

ausflora is currently not on CRAN. You can install its current developmental version using

```{r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("traitecoevo/ausflora")
```

Once its installed, we can load it into R using

```{r setup}
library(ausflora)
```

To demonstrate how to use ausflora, we will use an example dataset `gbif_lite` which is documented in ?gbif_lite

```{r example data}
dim(gbif_lite)

gbif_lite
```

## Retrieve taxonomic resources

The first step is to retrieve the entire APC and APNI name databases and store them locally as taxonomic resources. We achieve this using `load_taxonomic_resources()`.

There are two versions of the APC and APNI databases that you can retrieve with the `stable_or_current_data`argument. Calling:

-   `stable` will retrieve the recently archived version of the databases from our [GitHub releases](https://github.com/traitecoevo/ausflora/releases). This is set as the default option.
-   `current` will retrieve the up-to-date databases directly from the APC and APNI website.

Note that the databases are quite large so the initial retrieval of `stable` versions will take a few minutes. Once the taxonomic resources have been stored locally, subsequent retrievals will take less time. Retrieving `current` resources will always take longer since it is. Check out our [Resource Caching]() article to learn more about how the APC and APNIC databases are accessed, stored and retrieved.

```{r load resources}
# Benchmarking the retrieval of `stable` or `current` resources
stable_start_time <- Sys.time()
stable_resources <- load_taxonomic_resources(stable_or_current_data = "stable")
stable_end_time <-  Sys.time()

current_start_time <- Sys.time()
current_resources <- load_taxonomic_resources(stable_or_current_data = "current")
current_end_time <-  Sys.time()

# Compare times
stable_end_time - stable_start_time
current_end_time - current_start_time
```

For a more reproducible workflow, we recommend specifying the exact `stable` version you want to use.

```{r}
resources <- load_taxonomic_resources(stable_or_current_data = "stable", version = "0.0.2.9000")
```

## Standardise plant taxon names

Now we can query our taxonomic names against the taxonomic resources we just retrieved using `create_taxonomic_update_lookup()`. This all-in-one function will:

- Align your taxonomic names to APNI and APC using our [fuzzy matching algorithm]()
- Update taxonomic names to APC accepted name.

If you would like to learn more, take a look at the section "Closer look at name standardisation with 'ausflora'"

```{r}
library(dplyr)
library(tidyr)

update_gbif_names <- gbif_lite |> 
  drop_na(species) |> # This step is temporary bug
  pull(species) |> 
  create_taxonomic_update_lookup(resources = resources)

update_gbif_names
```

The `original_name` is the taxon name used in your original data.
The `aligned_name` is the taxon name we used to link with the APC to identify any synonyms. 
The `apc_name` is the currently, accepted taxon name used by the Australian Plant Census.

We recommend joining the output of `create_taxonomic_update_lookup()` back to your original data so all names

```{r}
gbif_lite |> 
left_join(update_gbif_names,  # Left joining updated taxon names back to original data 
          by = join_by(species == original_name)) |> 
  select(species, ends_with("name"), # Rearranging naming columns
         aligned_reason, infraspecificepithet:decimalLatitude)
```

## Plant established status across states/territories

'ausflora' can also provide the state/territory distribution for established status (native/introduced) from the APC.

We can access the established status data by state/territory using `create_species_state_origin_matrix()`

```{r}
# Retrieve status data by state/territory 
status_matrix <- create_species_state_origin_matrix(resources = resources)

status_matrix
```

Here is a breakdown of all possible values for `origin` 

```{r}
library(purrr)
library(janitor)

# Obtain unique values
status_matrix |> 
  select(-species) |> 
  flatten_chr() |> 
  tabyl()
```

The formal definitions of the various established status can be found at XX. 

You can get the breakdown of species by established status for a particular state/territory using `state_diversity_counts()`

```{r}
state_diversity_counts("NSW", resources = resources)
```

You can also quickly check if a plant taxa is a native using `native_anywhere_in_australia()`

```{r}
update_gbif_names |> 
  sample_n(1) |>  # Choosing a random species
  pull(apc_name) |> # Extracting this APC accepted name
  native_anywhere_in_australia(resources = resources)
```

## Closer look at name standardisation with 'ausflora'

`create_taxonomic_update_lookup` is a simple, wrapper, function for novice users that want to quickly check and standardise taxon names. For more experienced users, you can take a look at the sub functions `align_taxa()` and `update_taxonomy()` to see how taxon names are processed, aligned and updated.

### Aligning names to APNI and APC

This function will:

-   Clean up your taxonomic names <!-- (Getting rid of trailing whitespaces, case correction) -->
-   Align your taxonomic names using our [fuzzy matching algorithm]()

You can control the degree of fuzzy matching using the arguments `max_distance_abs` and `max_distance_rel`

<!-- Should the summary be updating during alignment?  -->

<!-- 0 names already matched; 0 names checked but without a match; 121 taxa yet to be checked -->

<!-- What do all these columns do? -->

<!-- Difference between cleaned and stripped and stripped name2 -->

```{r align names}
library(dplyr)
library(tidyr)

aligned_gbif_taxa <- gbif_lite |> 
  drop_na(species) |>  
  pull(species) |> 
  align_taxa(resources = resources)

aligned_gbif_taxa
```

For every `aligned_name`, `align_taxa()` will provide a `aligned_reason` which you can review as a table of counts:

```{r align reason}
library(janitor)

aligned_gbif_taxa |> 
  pull(aligned_reason) |> 
  tabyl() |> 
  tibble()
```

### Update taxonomic names to alignments

```{r update names}
updated_gbif_taxa <- aligned_gbif_taxa |> 
  pull(aligned_name) |> 
  update_taxonomy(resources = resources)
```


