---
title: "Introduction to ausflora"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ausflora}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When working with species data, it is important to verify your taxonomic names with an authoritative list and correct any out-of-date names. The ausflora package simplifies this process: 

- By accessing up-to-date taxonomic information from the [Australian Plant Census](https://biodiversity.org.au/nsl/services/search/taxonomy) and the [Australia Plant Name Index](https://biodiversity.org.au/nsl/services/search/names).
- 


ausflora is currently not on CRAN. You can install its current developmental version using

```{r install}
# install.packages("remotes")
remotes::install_github("traitecoevo/ausflora")
```

Once its installed, we can load it into R using

```{r setup}
library(ausflora)
```

To demonstrate how to use ausflora, we will use an example dataset `XX` which is documented in ?XX

```{r}

```

