---
title: "How to be more reproducible with APCalign"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

The article will show you how to use `APCalign` to update and align your plant taxonomic names in a more reproducible manner. The tips offered below will be particularly useful if you used our package and will share your code and data in your research paper or report. 

There are two components that we need to cited and their versions determined: 

- The `APCalign` package itself
- The taxonomic resources used by `APCalign` for aligning and updating your plant taxon names

Both of these components are updated for bug fixes, or to incorporate new taxonomic information and decisions. 

First let's load `APCalign`


```{r eval=TRUE}
library(APCalign)
```

#### APCalign R package version

To determine the version of the `APCalign` package itself: 

```{r}
packageVersion("APCalign")
```

#### Taxonomic Resources

`APCalign` allows users to load static downloads of taxonomic resources the APC and APNI or the latest version from the National Species List website. This functionality is specified using the `stable_or_current_data` argument of `load_taxonomic_resources()`. 

We recommend to always use `stable_or_current_data = "stable"` if you want your taxonomic alignment and update to be reproducible. The preset value is `stable_or_current_data = "stable"`. These static downloads are version controlled and stored in our repository as [releases](https://github.com/traitecoevo/APCalign/releases). 

```{r}
load_taxonomic_resources(stable_or_current_data = "stable")
```

By default, `load_taxonomic_resources()` will load the latest version of the static downloads.

```{r}
load_taxonomic_resources(
  stable_or_current_data = "stable",
  version = default_version()
)
```

In order to be more transparent, we recommend you to check what is the latest `default_version` before each alignment

```{r eval=TRUE}
default_version()
```

Then copying and pasting the output into `load_taxonomic_resources()` directly. This way makes the version of taxonomic resources more explicit in your code.

```{r}
resources <- load_taxonomic_resources(
  stable_or_current_data = "stable",
  version = "0.0.4.9000"
)
```

#### Citing the R package

For completion, you can also cite the R package by calling `citation()`. We also have a research article introducing the `APCalign`, we will share the details of its citation when it is in press. 

```{r, eval=TRUE}
citation("APCalign")
```

